---
title: "Reproducible Research: Peer Assessment 1"
author: "Jungmun Lee"
date: "26/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}
data <- read.csv("./activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
daily_steps <- aggregate(steps ~ date, data, sum)
```

```{r}
hist(daily_steps$steps, breaks=20, col="red",
    xlab="Steps per day",
    main="Histogram of the total number of steps")
```

# Calculate and report the mean and median of the total number of steps taken per day
    
```{r}
mean(daily_steps$steps)
median(daily_steps$steps)
```
    
## What is the average daily activity pattern?

Here's the sum of steps grouped by interval
```{r}
interval_steps <- aggregate(steps ~ interval, data, mean)
```

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
    
```{r}
plot(interval_steps$interval, interval_steps$steps, type = "l", 
     xlab = "Interval", ylab = "Steps", 
     main = "Average daily activity time series plot")
```
    
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval_steps[which.max(interval_steps$steps),]
```
    
## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data))
```
    
Misssing values will be replace with the mean for that 5-minute interval.
```{r}
na_data <- is.na(data$steps)
avg_interval <- tapply(data$steps, data$interval, mean, na.rm=TRUE)
data$steps[na_data] <- avg_interval[as.character(data$interval[na_data])]
```

Result after imputing missing values
```{r}
sum(is.na(data))
```
    
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
daily_steps <- aggregate(steps ~ date, data, sum)
```
    
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
hist(daily_steps$steps, breaks=20, col="red",
     xlab="Steps per day",
     main="Histogram of the total steps (missing values imputed)")
```
    
```{r}
mean(daily_steps$steps)
median(daily_steps$steps)
```
After imputing missing data the mean and the median of the total daily number of steps are the same.
    
## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
    
```{r}
data$daytype <- rep(NA, nrow(data))
data$daytype <- strftime(data$date, "%A")
data[data$daytype %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"),]$daytype <- "Weekday"
data[data$daytype %in% c("Saturday","Sunday"),]$daytype <- "Weekend"
```
    
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
daytype_steps <- aggregate(steps ~ interval + daytype, data, sum)

library(ggplot2)
ggplot(daytype_steps, aes(x=interval, y=steps, color=daytype)) +
  geom_line() +
  facet_wrap(~daytype, ncol=1, nrow=2)
```
